---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r global_options}
knitr::opts_chunk$set(fig.path='Figs/')
```

## Loading and preprocessing the data
```{r loading}
library(dplyr)
library(datasets)
library(ggplot2)
library(tibble)
library(lattice)

filename <- "activity.csv"
activity <- read.csv(filename)
dim(activity)
head(activity)
summary(activity)
```

## What is mean total number of steps taken per day?
```{r mean total and histogram}
steps_per_day <- tapply(activity$steps,activity$date, FUN=sum, rm.na = TRUE)

hist(steps_per_day, main = "Total number of steps taken per day and frequency", xlab = "Total steps by day")

steps_mean <- mean(steps_per_day, na.rm = TRUE)
steps_median <- median(steps_per_day, na.rm = TRUE)
cat("Mean of the total number of steps per day is ", steps_mean)
cat("Median of the total number of steps per day is ", steps_median)
```

## What is the average daily activity pattern?
```{r average daily activity}
avg_steps<-with(activity,tapply(steps,interval,mean,na.rm=TRUE))
avg_steps <- as.data.frame(avg_steps)


avg_steps <- avg_steps %>% rownames_to_column(var = "Interval")
plot(avg_steps$Interval, avg_steps$avg_steps, type = "l", main = "Averagy daily activity pattern")
```


## Imputing missing values
```{r missing values}
rownum <- which.max(avg_steps$avg_steps)
max <- avg_steps[rownum, "avg_steps"]
sum(is.na(activity$steps))
fill <- mean(avg_steps$avg_steps)
new_activity <- activity
count <- nrow(new_activity)
for(i in 1:count){
  if(is.na(new_activity$steps[i])) {
     new_activity$steps[i] <- fill
  }
}
```

```{r histogram and new mean/average}
new_steps_per_day <- tapply(new_activity$steps, new_activity$date, FUN=sum)

hist(new_steps_per_day, ylab = "Number of days", col = "Blue")
hist(steps_per_day, xlab = "Total steps by day", col = "Grey", add = T)
legend("topright", c("Imputed Data", "Non-NA Data"), fill=c("grey", "blue") )

steps_mean <- mean(new_steps_per_day, na.rm = TRUE)
steps_median <- median(steps_per_day, na.rm = TRUE)
cat("Mean of the total number of steps per day after filling NA with an average number of steps is ", steps_mean)
cat("Median of the total number of steps per day after filling NA with an average number of steps is ", steps_median)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r Weekdays and weekends}
new_activity$date <- as.Date(new_activity$date)
new_activity <- mutate(new_activity, Weekdays = weekdays(new_activity$date))


weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
new_activity$Weekdays <- c('weekend', 'weekday')[(weekdays(new_activity$date) %in% weekdays1)+1L]
```

```{r weekdays}
wkdays <- filter(new_activity, Weekdays == "weekday")
wkdays_avg_steps <- tapply(wkdays$steps, wkdays$interval, FUN=mean)
wkdays_avg_steps <- as.data.frame(wkdays_avg_steps)

wkdays_avg_steps <- wkdays_avg_steps %>% rownames_to_column(var = "Interval")
#plot(wkdays_avg_steps$Interval, wkdays_avg_steps$wkdays_avg_steps, type = "l", xlab = "Intervals", ylab = "Average Steps",main = "Weekdays")
wkdays_avg_steps <- mutate(wkdays_avg_steps, Week = "weekdays")

```

```{r weekends}
wkends <- filter(new_activity, Weekdays == "weekend")
wkends_avg_steps <- tapply(wkends$steps, wkends$interval, FUN=mean)
wkends_avg_steps <- as.data.frame(wkends_avg_steps)

wkends_avg_steps <- wkends_avg_steps %>% rownames_to_column(var = "Interval")
#plot(wkends_avg_steps$Interval, wkends_avg_steps$wkends_avg_steps, type = "l", xlab = "Intervals", ylab = "Average Steps",main = "Weekend")
wkends_avg_steps <- mutate(wkends_avg_steps, Week = "weekend")

colnames(wkdays_avg_steps) <- colnames(wkends_avg_steps) <- c("Interval", "Average_steps", "Week")
week_average <- rbind(wkdays_avg_steps, wkends_avg_steps)

week_average$Interval <- as.numeric(week_average$Interval)
week_average$Average_steps <- as.integer(week_average$Average_steps)
```
```{r week comparison plots}
xyplot(Average_steps~Interval|Week, type="l", data=week_average, layout = c(1,2), main = "Average steps per interval by day type")
```
install.packages(“knitr”)
library(“knitr”)
knit2html("/Users/masamiyamaguchi/Documents/GitHub/RepData_PeerAssessment1/PA1_template.Rmd", spin(knit = FALSE), force_v1 = TRUE)

